Puck Predictions: Unraveling the NHL Game Forecasting Riddle

### 2024-03-29 Latex/draft.tex
Added bayesian network citation and edited tsne graph.

\section{Data}
% Your content for this section goes here
Our data came from the hockeyR Github repository\cite{hockeyR-data}. This repository contains an abundance of data about every NHL game
that has occured since the 2010-11 season. This data includes information about the events that transpire in a game (hits, shots, goals, etc.),
which teams are playing, who is on the ice, and the final score of the game. The data is stored in a series of {\tt .csv.gz} files, allowing for
easy access and manipulation.

@@

The computational complexity of Bayesian Network inference is high, with exact inference being an NP-hard problem TODO CITE. 
Using the python package pgmpy, we originally tried to fit a network with all 26 of our features, but our computational resources failed to fit this network.
Then, to get a baseline for our future predictions, we simply fitted the model with the base features of time remaining (tr), home goals (hg), away goals (ag), home shots (hs), away shots (as), home blocked shots (hbs), and away blocked shots (abs). 
These features were chosen as priors because of our opinion that they are the most important to the game, based upon our knowledge of hockey.

@@

\subsection{t-SNE, UMAP, and PCA}

@@

To evaluate our simulation's effectiveness at accurately modeling actual hockey 
games, we combined a dataset with the final event counts for 1500 actual hockey 
games and for 1500 simulated games. We then performed PCA, t-SNE and UMAP 
dimensionality reductions using various perplexity and neighbor hyperparameters 
to see if these algorithms clustered the synthetic games and actual games in 
different clusters. As shown below in \ref{fig:simulation_v_actual}, the simulated
games and actual games are all clustered together, thus demonstrating that our
simulation effectively emulates live NHL games.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{images/simulation_v_actual_dimred.png}
    \caption{PCA, t-SNE and UMAP performed on a dataset with final event counts for 1500 simulated and 1500 actual NHL games. The joined grouping demonstrates that our simulation accurately emulates live NHL games}
    \label{fig:simulation_v_actual}
\end{figure}

>>>>

\section{Data}
% Your content for this section goes here
Our data came from the hockeyR Github repository~\cite{hockeyR-data}. This repository contains an abundance of data about every NHL game
that has occured since the 2010-11 season. This data includes information about the events that transpire in a game (hits, shots, goals, etc.),
which teams are playing, who is on the ice, and the final score of the game. The data is stored in a series of {\tt .csv.gz} files, allowing for
easy access and manipulation.

@@

The computational complexity of Bayesian Network inference is high, with exact inference being an NP-hard problem~\cite{pmlr-vR0-chickering95a}. 
Using the python package pgmpy, we originally tried to fit a network with all 26 of our features, but our computational resources failed to fit this network.
Then, to get a baseline for our future predictions, we simply fitted the model with the base features of time remaining (tr), home goals (hg), away goals (ag), home shots (hs), away shots (as), home blocked shots (hbs), and away blocked shots (abs). 
These features were chosen as priors because of our opinion that they are the most important to the game, based upon our knowledge of hockey.

@@

@@

To evaluate our simulation's effectiveness at accurately modeling actual hockey 
games, we combined a dataset with the final event counts for 1500 actual hockey 
games and for 1500 simulated games. We then performed PCA, t-SNE and UMAP 
dimensionality reductions using various perplexity and neighbor hyperparameters 
to see if these algorithms clustered the synthetic games and actual games in 
different clusters. As shown below in~\ref{fig:simulation_v_actual}, the simulated
games and actual games are all clustered together, thus demonstrating that our
simulation effectively emulates live NHL games.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{images/pca_tsne_umap_sim_act.png}
    \caption{PCA, t-SNE, and UMAP performed on a dataset with final event counts for 1500 simulated and 1289 actual NHL games. The joined grouping demonstrates that our simulation accurately emulates live NHL games}
    \label{fig:simulation_v_actual}
\end{figure}
